Viscous Transport in Eroding Porous Media

### 2019-08-05 docs/transport/erosion_transport.tex
add new figs in transport paper
small changes to paper

One way to define the anomalous diffusion rate is by using tracer
trajectories that are governed by equation~\eqref{eqn:trajectories}.
Given a set of trajectories, we define $\lambda_j(t)$ to be the distance
travelled by the particle
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss_j(\tau)\|\, d\tau, 
    \quad j=1,\ldots,N_p.
\end{align}
Having formed $N_p$ trajectories, the first and second ensemble moments
are
\begin{align}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2.
\end{align}
At early times, the particles have not explored much of the geometry,
and we expect a ballistic motion $\sigma_\lambda^2 \sim t^2$.  However,
as the particles pass the grains, their trajectories are altered, and we
expect that $\sigma_\lambda^2 \sim t^\alpha$, with $\alpha < 1$, which
would indicate that the flow is sub-diffusive.  With respect to erosion,
as the bodies erode, we expect that the channelization of the geometry
results in more trajectories undergoing a ballistic motion, and this
would indicate a larger anomalous diffusion rate.

@@

\begin{figure}[H]
\center
\includegraphics*[width =0.5\linewidth]{./figs/tort_eulerian}
\caption{\label{fig:Eroding20tort_all} The Eulerian and Lagrangian approach of tortuosity. 
The blue dots are Eulerian approach of tortuosity and the red stars and 
square are Lagrangian approach of tortuosity.}
\end{figure}
%\begin{figure}[H]
%\center
%\includegraphics*[width =0.5\linewidth]{./figs/tort_diff_top10_268}
%\caption{\label{fig:Eroding20tort_traj} The largest 10 difference of 
%local tortuosity $\tau$ at two neighboring tracers in
% 20 nearly touching bodies eroding when the porosity is 62.9 \% 
%and the tortuosity is 1.0979. The same color of a pair of trajectories 
%means they are from two neighboring tracers.}
%\end{figure}

\begin{figure}[H]
\center
\includegraphics*[width =0.5\linewidth]{./figs/anomalous_diffusion20d_new}
\caption{\label{fig:Eroding20anomalous} The analysis of anomalous diffusion in
 20 nearly touching bodies with respect to six porosities.}
\end{figure}

>>>>

One way to define the anomalous diffusion rate is by using tracer
trajectories that are governed by equation~\eqref{eqn:trajectories}.
Given a set of trajectories, we define $\lambda_j(t)$ to be the distance
travelled by the particle
\begin{align}
  \lambda_j(t) = \int_{0}^t \|\ss'_j(\tau)\|\, d\tau, 
    \quad j=1,\ldots,N_p.
\end{align}
Having formed $N_p$ trajectories, the first and second ensemble moments
are
\begin{align}
  \langle \lambda \rangle (t) = 
    \frac{1}{N_p} \sum_{j=1}^{N_p} \lambda_j(t), \quad 
    \sigma_\lambda^2(t) = \frac{1}{N_p} 
    \left[\lambda_j(t) - \langle \lambda \rangle(t) \right]^2.
\end{align}
At early times, the particles have not explored much of the geometry,
and we expect a ballistic motion $\sigma_\lambda^2 \sim t^2$.  However,
as the particles pass the grains, their trajectories are altered, and we
expect that $\sigma_\lambda^2 \sim t^\alpha$, with $\alpha < 1$, which
would indicate that the flow is sub-diffusive.  With respect to erosion,
as the bodies erode, we expect that the channelization of the geometry
results in more trajectories undergoing a ballistic motion, and this
would indicate a larger anomalous diffusion rate.

Towards the end of the erosion simulation, the porosity is low, and we
expect that the ensemble moments behave similarly to transport in a
pipe.  In a grainless geometry, the trajectory originating at
$(x_0,y_0)$ is
\begin{align}
  s_1(t) &= (1-y_0^2)t + x_0, \\
  s_2(t) &= y_0.
\end{align}
Averaging over $y_0 \in [-1,1]$, the first and second ensemble moments
are
\begin{align}
  \langle \lambda \rangle (t) = \frac{2}{3}t, \quad 
    \sigma_\lambda^2(t) = \frac{4}{45}t^2.
\end{align}

@@

In Figure~\ref{fig:Eroding20tort_all}, we plot the tortuosity as a
function of the porosity from the initial porosity of $0.38$ until all
the bodies have eroded.  The tortuosity is computed with both the
length of the Lagrangian trackers~\eqref{eqn:tortuosity1} and using the
spatial average of the velocity on an Eulerian
grid~\eqref{eqn:tortuosity2}.  The red square mark corresponds to the
porosity in Figure~\ref{fig:Eroding20tort}(c).  The two methods of
calculating the tortuosity give similar results with any discrepancy
being accounted for by slow regions of recirculation.  Computing the
velocity on an Eulerian grid is computationally easier than simulating
many trajectories, so we use the Eulerian approach for all future
examples.  As the bodies erode, this creates wide channels where
trajectories undergo only minor vertical deflections, and this explains
why the tortuosity eventually decreases with porosity.  However, at the
early stages of the erosion, the tortuosity increases with porosity,
which can be explain as follows.   In the absence of erosion
(left plot in Figure~\ref{fig:Eroding20vort}), many of the trajectories,
such as those initialized between bodies 15 \& 18, only perform minor
deflections to pass through the narrow regions, albeit, very slowly.
However, as erosion starts to open the channels, the trajectories
deflect into the fast regions, and this increases the amount of vertical
deflection, and therefore the tortuosity.
\todo[inline]{BQ: Need to work on this paragraph}

\begin{figure}[H]
\center
\includegraphics*[width =0.5\linewidth]{./figs/tort_eulerian}
\caption{\label{fig:Eroding20tort_all} The tortuosity inside of an
eroding geometry initialized with 20 bodies.  The tortuosity is
calculated using the Eulerian method~\eqref{eqn:tortuosity2} (blue dots)
and Lagrangian method~\eqref{eqn:tortuosity1} (red stars).  The red
square correspond to the geometry in Figure~\ref{fig:Eroding20tort}(c).}
\end{figure}

\begin{figure}[H]
\center
\includegraphics*[width =0.5\linewidth]{./figs/anomalous_diffusion20d_new}
%\caption{\label{fig:Eroding20anomalous} The
%mean squared displacement of the tracer trajectories  at several
%different stages of the erosion process.  As the porosity increases, the
%mean squared displacement tends towards $\sigma^2_\lambda(t) \sim t^$
%}
\end{figure}
